Offshore Crabs/Hectare Modeling

### 2020-01-07 2019_edits/R/offshore_analyses/crabs_per_hectare.Rmd
nb model with offset

## Tweedie Models {.tabset .tabset-pills}

Tweedie distributions are similar to a negative-binomial in their application for datasets that exhibit zero inflation. Tweedie models are applicable in cases where the response is continuous i.e. catch/area.

@@

```{r}
seamap <- seamap %>% rename_all(tolower)
seamap_west <- seamap %>% 
  filter(statzone != 12,
         statzone %in% 11:21)

mod1a <- glm(cpue_towspd ~ year_f + region + season + survey_design +  #Additive covariates
            year_f * region,                                           #Interaction terms
            family = tweedie(),
            data = seamap_west)

# summary(mod1a)
# 
# emmeans::emmip(mod1a, region ~ year_f) 
```

@@

```{r}
broom::glance(mod1a) %>% kable()
```

@@

```{r}
#Predicted abundance dataframe
mod1a.dat <- emmeans(mod1a, 
                        specs = "year_f" , 
                        by    = "region", 
                        type  = "response") %>% 
  as.data.frame()

#Try this
emm <- update(emmeans(m1, "am"), tran = "log")
summary(emm, type = "response")

#Standardize to mean
mean_pred <- mean(mod1a.dat$response)
mod1a.dat <- mod1a.dat %>% mutate(Std_Index = response/mean_pred,
                                CV = SE/response,
                                LCI = asymp.LCL/mean_pred,
                                UCI = asymp.UCL/mean_pred,
                                Year_n = as.numeric(as.character(year_f)))

#make scale_x_discrete breaks b/c ggplot is dumb
ystart <- min(as.numeric(seamap$survey_year))
yend   <- max(as.numeric(seamap$survey_year))
ynames <- lubridate::year(seq.Date(as.Date(paste(ystart), "%Y"), 
                   as.Date(paste(yend), "%Y"), 
                   by = "year"))

#ggplot of standardized index of abundance
p <- mod1a.dat %>% 
  ggplot(aes(year_f, response)) +
      geom_ribbon(aes(Year_n, ymin = asymp.LCL, ymax = asymp.UCL), fill = "gray80") +
      geom_line(aes(Year_n, response, group = 1)) +
      geom_point(aes(Year_n, response)) +
      geom_hline(aes(yintercept = 0), color = "darkred", linetype = 2) +
      scale_x_discrete(limits = c(as.numeric(ystart:yend)), labels = ynames) +
      xlab("") + ylab("Crabs per Hectare") +
      theme(axis.text.x=element_text(angle=45, hjust = 1, vjust=1), legend.position = "bottom") +
      facet_wrap(~region)
p

@@

## Negative Binomial Models {.tabset .tabset-fade .tabset-pills}

>>>>

## NEgbin w/ Offset {.tabset .tabset-pills}

@@

```{r}
seamap <- seamap %>% rename_all(tolower)
seamap_west <- seamap %>% 
  filter(statzone != 12,
         statzone %in% 11:21) %>% 
  mutate(statzone = factor(statzone))

mod1a <- MASS:::glm.nb(sapidus_catch ~ year_f + region + season +   #Additive covariates
            year_f * region + 
            offset(log(area_hectares)),
            data = seamap_west)

summary(mod1a)

emmeans::emmip(mod1a, region ~ year_f)
```

@@

```{r}
broom::glance(mod1a) %>% kable()

#Percent deviance explained
(mod1a$null.deviance - mod1a$deviance) / mod1a$null.deviance
```

@@

#Estimated Marginal Means with an offset
#parameter in the call of offset = 0:
mod1a.dat <- as.data.frame(emmeans(mod1a, 
                                  specs = "year_f", 
                                  by    = "region",
                                  type  = "response", 
                                  offset = 0))
mod1a.dat$CV <- nbemeans$SE/nbemeans$response

#make scale_x_discrete breaks b/c ggplot is dumb
ystart <- min(as.numeric(seamap_west$survey_year))
yend   <- max(as.numeric(seamap_west$survey_year))
ynames <- lubridate::year(seq.Date(as.Date(paste(ystart), "%Y"), 
                   as.Date(paste(yend), "%Y"), 
                   by = "year"))

#ggplot of standardized index of abundance
p <- mod1a.dat %>% 
  mutate(year_n = as.numeric(as.character(year_f))) %>% 
  ggplot(aes(year_n, response)) +
      geom_ribbon(aes(year_n, ymin = asymp.LCL, ymax = asymp.UCL), fill = "gray80") +
      geom_line(aes(year_n, response, group = 1)) +
      geom_point(aes(year_n, response)) +
      geom_hline(aes(yintercept = mean(seamap_west$cpue_towspd)), color = "darkred", linetype = 2) +
      scale_x_discrete(limits = c(as.numeric(ystart:yend)), labels = ynames) +
      labs(
        x = NULL, 
        y = "Crabs per Hectare",
        caption = "Red dashed line is the grand mean of crabs/hectare, representing the null model.") +
      theme(axis.text.x=element_text(angle=45, hjust = 1, vjust=1), legend.position = "bottom") +
      facet_wrap(~region)
p

@@

### Observed vs. Predicted

Model prediction (black) & observed mean values (blue), both standardized by overall mean

```{r}
#Observed data
mean_obs <- mean(seamap$Sapidus_Catch)
obs_timeline <- seamap %>% 
  group_by(Survey_Year) %>% 
  summarise(Nobs = n(),
            PercentPos = mean(Crab_presence),
            Mean_Catch = mean(Sapidus_Catch)) %>% 
  mutate(Index_Obs = Mean_Catch/mean_obs)

#Plot again with observed data set on top in blue 
p + geom_line(data = obs_timeline, 
              aes(Survey_Year, Index_Obs, group = 1), 
              col = "royalblue", linetype = 2) +
    geom_point(data = obs_timeline, 
               aes(Survey_Year, Index_Obs), col = "royalblue") +
  labs(title = "Model Estimates and Yearly observed means")
```
